Refining from the BON High-level Specification Language to the VHDL-ASM Hardware Description Language

### 2009-07-27 papers/BON_VHDL/paper.tex
documents section start

\newcommand{\tablesize}{\footnotesize}
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\myhref}[2]{\ifpdf\href{#1}{#2}\else\htmladdnormallinkfoot{#2}{#1}\fi}
%\newcommand{\myhref}[2]{\emph{#2}}
\newcommand{\todo}{\textbf{TODO:}}
\newcommand{\ST}{\emph{SenseTile}\xspace}
\newcommand{\STs}{\emph{SenseTiles}\xspace}
\newcommand{\STS}{\emph{SenseTile Simulator}\xspace}
\newcommand{\datastore}{\STs Scientific Datastore\xspace}
\newcommand{\computefarm}{The \STs Scientific Compute Farm\xspace}
\newcommand{\computefarmlong}{UCD CASL \STs Software and Data Compute Server Farm\xspace}
\newcommand{\sensorfarm}{The \STs\xspace}

%---------------------------------------------------------------------
% New commands, macros, \etc
%---------------------------------------------------------------------

@@

\title{Refining from the BON High-level Specification Language to the
  VHDL Hardware Description Language}

@@

The \STs have several unusual built-in sensors and emitters, sensor
nodes can be extended via a general-purpose USB bus, and \STs contain
an FPGA that is dynamically updated on a per-experiment basis.  \STs
are custom-designed in VHDL.  Like many hardware-based systems, to
write new software for or against the \ST architecture one either uses
the actual hardware or one runs a \STS.  But this simulator emulates a
piece of hardware, one would like to know if the simulator
\emph{actually} simulates the hardware.  We have developed an approach
through which we refine the VHDL specification up to the formal BON
specification language, and from this specification we generate an
executable, functional model-based specification in the JML language.
We have used this specification to implement a fully validated and
verified \ST simulator.

@@

%======================================================================
\section{Introduction}

An introduction.

>>>>

\newcommand{\tablesize}{\footnotesize}
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\myhref}[2]{\ifpdf\href{#1}{#2}\else\htmladdnormallinkfoot{#2}{#1}\fi}
%\newcommand{\myhref}[2]{\emph{#2}}
\newcommand{\todo}{\textbf{TODO:}}
\newcommand{\ST}{\emph{SenseTile}\xspace}
\newcommand{\STs}{\emph{SenseTiles}\xspace}
\newcommand{\STS}{\emph{SenseTile Simulator}\xspace}
\newcommand{\datastore}{\STs Scientific Datastore\xspace}
\newcommand{\computefarm}{The \STs Scientific Compute Farm\xspace}
\newcommand{\computefarmlong}{UCD CASL \STs Software and Data Compute Server Farm\xspace}
\newcommand{\sensorfarm}{The \STs\xspace}
\newcommand{\Sc}{\emph{system\_chart}\xspace}
\newcommand{\Clc}{\emph{cluster\_chart}\xspace}
\newcommand{\Csc}{\emph{class\_chart}\xspace}
\newcommand{\Ft}{\emph{feature}\xspace}
\newcommand{\ent}{\emph{entity}\xspace}
\newcommand{\arch}{\emph{architecture}\xspace}
%---------------------------------------------------------------------
% New commands, macros, \etc
%---------------------------------------------------------------------

@@

\title{Refining from the BON High-level Specification Language to the
  VHDL-ASM Hardware Description Language}

@@

The \STs, as an embedded system, have several unusual built-in sensors 
and emitters, sensor nodes can be extended via a general-purpose USB bus, 
and \STs contain an FPGA that is dynamically updated on a per-experiment 
basis.  \STs are custom-designed in VHDL and VHDL-ASM.  Like many hardware-based systems,
 to write new software for or against the \ST architecture one either uses
the actual hardware or one runs a \STS.  But this simulator emulates a
piece of hardware, one would like to know if the simulator
\emph{actually} simulates the hardware.  We have developed an approach
through which we refine the VHDL specification up to the formal BON
specification language, and from this specification we generate an
executable, functional model-based specification in the JML language.
We have used this specification to implement a fully validated and
verified \ST simulator.

@@

%======================================================================
\section{Introduction}
The \STs, represents heterogeneous collection of digital, analog 
and mixed signal hardware components. Embedded system complexity 
requires the need to strive for formal descriptions each component 
separatelly in order to produce specification of the \STs simulator. 
The development of complex system requires sophisticated tool and methodology 
in order to be able to complete a correct design in time.

The authors will illustrate in this paper how BON can represent a high 
level analogue mapping and will further illustrate a procedure for 
transformation from BON to VHDL-ASM. As an example, 
we will give practical aspect of refinement from BON to VHDL-ASM,
applied on A/D convertor, in order to produce formal specification 
and software simulation.
%=====================================================================

\section{VHDL-ASM}
\subsection{Hardware description language}

\section{Choice of BON as the modeling tool}
%Brief description and motivation.
The Business Object Notation (BON) is an analysis and design notation for 
object-oriented systems  which emphasizes \emph{seamlessness}, \emph{reversibility} and software \emph{contracting}.
Let us define these principles precisely and point out why the principles are important in VHDL-ASM simulation.

\begin{itemize}
\item \emph{Seamlessness} is the use of a continuous process throughout the software lifecycle.
\item \emph{Reversibility} is the support for both forward and backward development process: from analysis to
design and implementation, and back.
In VHDL-AMS, modelling represents an continious top-down-bottom-up process. Top-down approach serves for design whereas bottom-up
serves for verification.
The possibility of seamlessly translating changes made during a certain development phase back into earlier phases,
facilitates a stable and predictable environment and ensures the various models consistent with the code.

\item \emph{Contracting} is the precise definition, for each software element, of the associated semantic properties. 
BON use a full assertion mechanism, allowing analysts to specify not only the structure of a system but also its
semantics ( constraints and invariants of the expected results ).VHDL-AMS provides strongly typed interface constructs 
that ensure clarity in the interface specifications.

\end{itemize}

%\todo{dragan 25/07/2009 - Figure out relation: Seamlessness and VHDL-ASM}
%\todo{dragan 25/07/2009 - Figure out relation: Reversibility and VHDL-ASM. Why is Reversibility important in VHDL design.}
Therefore as an approach of best modelling practices, and as the language based on contract design 
and refinement calculus, BON is the obvious choice for the our simulator.

\subsection{Mapping between BON and VHDL-ASM}
%\todo{dragan 25/07/2009 - Figure out : Mapping from BON  to entity/architecture structure in VHDL-ASM.}
\begin{itemize}
\item \Sc: can be represented with high abstract level of the topmost entity in a design hierarchy.
\item \Clc: to be described.
\item \Csc: can be equated with the primary unit(\ent) and secondary unit (\arch) structure in VHDL-AMS.
The \Csc's name corresponds to the name of the entity in VHDL-AMS.
\item \Ft: The feature section of the \Csc, where the mathematical transfer function of the component
is entered, holds the same information as the \arch section of the VHDL-AMS.
\end{itemize}
\subsection{Refinement from BON to VHDL-ASM}
